---
title: "Plot Two variables"
author: "Chiara"
date: "October 4, 2015"
output: html_document
---

#Data Analysis with R
##Udacity course
[Click here](https://www.udacity.com/me#!/) for Udacity courses 

[GGplot 2 Geoms](http://docs.ggplot2.org/current/)

[Documentation for smooths](http://www.ats.ucla.edu/stat/r/faq/smooths.htm)

[Color ggplot](http://www.cookbook-r.com/Graphs/Colors_%28ggplot2%29/)

Continuing from "Plot pseudoFacebook"
```{r}
pf = read.csv("C:/Users/cdiloreto/Google Drive/Programming/Data sets and exercises/pseudo_facebook.tsv", sep = "\t")

# Load ggplot2
library(ggplot2)
```

Plot 2 Variables, `age` and `friend_count`
```{r}
qplot(x = age, y = friend_count, data = pf)
```

Ggplot syntax
```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) + 
    geom_point(alpha = 1/20) +
    xlim(13,100)
```

Use geom_jitter instead of geom_point
```{r ggplot jitter}
ggplot(aes(x = age, y = friend_count), data = pf) + 
    geom_jitter(alpha = 1/20) +
    xlim(13,100)
```

#####Exercise: Use `coord_trans` to transform the y-axis
```{r ggplot coordinates transform}
ggplot(aes(x = age, y = friend_count), data = pf) + 
    geom_jitter(alpha = 1/20) +
    xlim(13,100) +
  coord_trans(limy = c(0, 1000))
```

Solution:
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + 
    geom_point(alpha = 1/20) + ## geom_jitter here produces an error
    xlim(13,100) +
  coord_trans(y = "sqrt")

## Syntax for geom_jitter
ggplot(aes(x = age, y = friend_count), data = pf) + 
    geom_jitter(alpha = 1/20, position = position_jitter(h = 0)) + 
    xlim(13,100) +
  coord_trans(y = "sqrt")

```

Exercise: Investigate the friendships_initiated vs. age
```{r}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point()

# Geom jitter
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_jitter()

# add alpha value
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_jitter(alpha = 1/10)
```

Solution:
```{r}
# same thing as above - different syntax
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 1/10, position = "jitter")

# transform the y-axis and change syntax to avoid errors
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_jitter(alpha = 1/10, position = position_jitter(h = 0)) +
  coord_trans(y = "sqrt")

```

###Transform data frames
####Dplyr package
```{r}
library(dplyr)

# Use group_by() to find age groups
age_groups = group_by(pf, age)

# Then summarise function to find the mean and the median of friend_count by age
pf.friend_count_by_age = summarise(age_groups, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) # this paramenter can only be used with summarise function to count number of occurences

# Investigate the new data frame
head(pf.friend_count_by_age)

```

Alternate code
```{r}
pf.friend_count_by_age = pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)
```

Exercise: plot the friend_count_mean vs. age
```{r}
ggplot(aes(x = age, y = friend_count_mean), data = pf.friend_count_by_age) + geom_line()
```

### Overlaying Summaries with Raw Data

```{r Overlaying Summaries with Raw Data}
# Little try myself
ggplot(aes(x = age, y = friend_count_mean), data = pf.friend_count_by_age) + geom_line() +
  geom_point()

# Lesson: copy code from above, raw data
ggplot(aes(x = age, y = friend_count), data = pf) + 
    geom_jitter(alpha = 1/20, position = position_jitter(h = 0)) + 
    xlim(13,100) +
  coord_trans(y = "sqrt")

# Overlay the mean values
ggplot(aes(x = age, y = friend_count), data = pf) + 
    xlim(13,90) +
    geom_jitter(alpha = 1/20, position = position_jitter(h = 0),
                color = "orange" ) + 
    coord_trans(y = "sqrt") +
    geom_line(stat = "summary", fun.y = mean) +
    geom_line(stat = "summary", fun.y = median,
              linetype = 3, color = "red") +
    geom_line(stat = "summary", fun.y = quantile, probs = 0.9,
            linetype = 2, color = "blue") +
  theme_bw()


# Change coord_cartesian values to "zoom in"
ggplot(aes(x = age, y = friend_count), data = pf) + 
    coord_cartesian(xlim = c(13,70), ylim = c(0,1000)) +
    geom_jitter(alpha = 1/20, position = position_jitter(h = 0),
                color = "orange" ) + 
    geom_line(stat = "summary", fun.y = mean) +
    geom_line(stat = "summary", fun.y = median,
              linetype = 2, color = "red") +
    geom_line(stat = "summary", fun.y = quantile, probs = 0.9,
            linetype = 2, color = "blue") +
  theme_bw()


```

###Correlation
```{r}
?cor.test
```

Exercise:
```{r}
##Find correlation between age and friend_count

cor.test(pf$age, pf$friend_count)
```

Solution:
```{r}
# same as above can be done like this
with(pf, cor.test(age, friend_count))
```

Find correlation between age and friend count for users that have less than 70 years. 
Values above 70 y-o are likely to be incorrect (fake accounts).
```{r}
with(subset(pf, age <= 70), cor.test(age, friend_count))

```

Exercise:
```{r}
# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x).

max(pf$likes_received)
max(pf$www_likes_received)

ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
    geom_point() +
    coord_cartesian(xlim = c(0,15000), ylim = c(0,15000))

```

Solution:
```{r}
# Plot based on the 95% of the observations (to exclude outliers)
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
    geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95))

# add geom_smooth curve
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
    geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
    geom_smooth(method = "lm", color = "red")

# ?geom_smooth
# lm = linear mode
```

Exercise: find correlation between www_likes_received and likes_received
```{r}
cor.test(pf$www_likes_received, pf$likes_received)
cor.test(pf$www_likes_received, pf$likes_received, conf.level = 1)
```


Exercise: dataframe Mitchell
```{r}
library(alr3)
data("Mitchell")
head(Mitchell)
range(Mitchell$Month)

## Line plot
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
    geom_line()

## Correlation
cor.test(Mitchell$Month, Mitchell$Temp)


## Add a scale to x to see variance by month
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
    geom_line() +
  scale_x_discrete(breaks = seq(0,203,12))

## use %% operator (modulus)
ggplot(aes(x = Month%%12, y = Temp), data = Mitchell) +
    geom_point() +
  scale_x_discrete(breaks = seq(0,203,12))

```

Back to friend_count_mean
```{r}
ggplot(aes(x = age, y = friend_count_mean), data = pf.friend_count_by_age) + geom_line()

pf.friend_count_by_age[17:19,]

```

Exercise: create age_with_months
```{r}
pf$age_with_months = NA

View(pf)

pf$age_with_months = pf$age + (1 - pf$dob_month/12)

## Solution
pf$age_with_months = pf$age + (12 - pf$dob_month)/12

```

```{r}
# Create a new data frame called
# pf.fc_by_age_months that contains
# the mean friend count, the median friend
# count, and the number of users in each
# group of age_with_months. The rows of the
# data framed should be arranged in increasing
# order by the age_with_months variable.

# For example, the first two rows of the resulting
# data frame would look something like...

# age_with_months  friend_count_mean	friend_count_median	n
#              13            275.0000                   275 2
#        13.25000            133.2000                   101 11


# See the Instructor Notes for two hints if you get stuck.
# This programming assignment will automatically be graded.

# DO NOT DELETE THESE NEXT THREE LINES OF CODE
# ========================================================================
#pf <- read.delim('/datasets/ud651/pseudo_facebook.tsv')
#pf$age_with_months <-pf$age + (1 - pf$dob_month / 12)
#suppressMessages(library(dplyr))

# ENTER YOUR CODE BELOW THIS LINE
# ========================================================================

# Use group_by() to order by age groups by month
age_groups_with_month = group_by(pf, age_with_months)

# Then summarise function to find the mean and the median of friend_count by age_with_month
pf.fc_by_age_months = summarise(age_groups_with_month, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) # this paramenter can only be used with summarise function to count number of occurences

# Investigate
head(pf.fc_by_age_months)

# ggplot
ggplot(aes(x = age_with_months, y = friend_count_mean), data = pf.fc_by_age_months) +
  geom_point()

### Try to do it with the alternate code

```

Now join side by side the two graphs, by age and by age_with_month
```{r}
## Subset the data to users younger than 70
p1 = ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line() +
  geom_smooth()

## Subset the other graph with same criteria
p2 = ggplot(aes(x = age, y = friend_count_mean), 
       data = subset(pf.friend_count_by_age, age < 71)) + 
      geom_line() +
  geom_smooth()

## Create a third plot
p3 = ggplot(aes(x = round(age/5)*5, y = friend_count),
            data = subset(pf, age < 71)) +
      geom_line(stat = "summary", fun.y = mean)

##
library(gridExtra)

grid.arrange(p1, p2, p3, ncol = 1)
```

Check [documentation for smooths](http://www.ats.ucla.edu/stat/r/faq/smooths.htm)
